---
title: "eda"
author: "Divya Bisht"
date: "10/4/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r data_import, cache = TRUE}
library(rnoaa)
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
     tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())

```

```{r}
weather_df %>%
  group_by(name, month)
```

How many unique variables are there?
-Start with summarize function

count first
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(number = n())


weather_df %>%
  group_by(month) %>%
  summarize(number = n())
```

## Grouped summaries
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(number = n(), 
      mean_tmax = mean(tmax))
```
Mean function has: 
-na.rm = should I remove missing values? Set to false by default

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(number = n(), 
      mean_tmax = mean(tmax, na.rm = TRUE))
```
Median
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(number = n(),
      median_precip = median(prcp, na.rm = TRUE), 
      sd_tmin = sd(tmin, na.rm = TRUE))
```

Still a tibble
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(number = n(),
      mean_tmax = mean(tmax, na.rm = TRUE), 
      sd_tmin = sd(tmin, na.rm = TRUE)) %>%

ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point()
```
Month data frame
```{r}
month_df = weather_df %>%
  group_by(name, month) %>%
  summarize(number = n(),
      mean_tmax = mean(tmax, na.rm = TRUE),
      median_precip = median(prcp, na.rm = TRUE),
      sd_tmin = sd(tmin, na.rm = TRUE)) %>%
  
  ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
    geom_point(),
    geom_line(data = month_df, aes(x = month, y = mean_tmax))
```
Tidy is not always best for people

```{r}
weather_df %>%
   group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>%
      spread(key = month, value = mean_tmax) %>%
  knitr::kable(digits = 3)
```

##Grouped mutate
  
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE))

```

Another example
-going to take the average temp in each place and repeat the average temp of that please in that column

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE), 
         centered_tmax = tmax - mean_tmax) %>%
  
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point()
```

##Window functions 
Use ranks
-tells you the ordering of values. Like tell you the coldest/hottest day in a certain month
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(tmax)) %>%
  filter(tmax_rank < 2) #this tells you each instance there was the coldest day.
```
if we want to get warmest day
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(desc(tmax))) %>%
  filter(tmax_rank < 2)
```

How much did observations change in certain variables(maybe over time?) 
use the lag function. 
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(lag_tmax = lag(tmax))
```
How neat is this? 
  -how much variability is there in the day-to-day temp of the locations?
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(lag_tmax = lag(tmax), 
         one_day_change = tmax - lag_tmax) %>%
  summarize(sd_one_day_change = sd(one_day_change, na.rm = TRUE))
```



